Rust for Operating System Development: A Systematic Literature Review

### 2024-04-10 paper.tex
Working on slr paper

\documentclass[sigconf]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\pagestyle{plain} %to show page numbers

@@

\begin{document}
\title{Memory Safety: A Systematic Literature Review}

\author{Shane Panter}
    \affiliation{%
    \institution{Boise State University}
    \city{Boise}
    \state{Idaho}
    \country{USA}}
    \email{shanepanter@boisestate.edu}

\author{Nasir U. Eisty}
    \affiliation{%
    \institution{Boise State University}
    \city{Boise}
    \state{Idaho}
    \country{USA}}
    \email{nasireisty@boisestate.edu}

@@

The migration of legacy C codebases to memory-safe programming languages like Rust is of paramount
importance in modern software development, aiming to mitigate common security vulnerabilities and
memory-related errors. This paper presents a systematic literature review (SLR) focusing on
strategies and methodologies for converting C code to Rust while minimizing the utilization of
unsafe blocks. Our study aims to provide a comprehensive overview of existing research, identify
gaps, and suggest future directions in this domain. Through a rigorous search process, we
synthesized relevant studies and extracted key findings to offer insights into effective approaches
for ensuring memory safety during code migration.

@@

\subsubsection{Research Questions}
We have defined the following research questions to drive this study.

\begin{itemize}

    \item \textbf{RQ1:} What are the existing methodologies and strategies for converting C code to
      Rust?
    \item \textbf{RQ2:} How do these methodologies address memory safety concerns and minimize the
      use of \lstinline{unsafe} blocks?
    \item \textbf{RQ3:} What are the challenges and limitations encountered in the conversion process?
\end{itemize}

@@

We employed a robust multi-step search strategy across three digital databases, ACM Digital Library,
IEEE Xplore, and Google Scholar in order to find all the current research that has been done. These
databases cover a wide range of current software engineering and computer science literature. We
leveraged the advanced search features of both the ACM Digital Library and IEEE Xplore to search the
titles and abstracts for keywords using boolean search operators. We then searched Google Scholar
using the same keywords and operators sorted by relevance and took the top 250 results.

@@

The search criteria for each database is shown in table~\ref{tab:keywords}.  First, we searched
titles and abstracts in the ACM Digital Library and IEEE Xplore with the keywords ``C rust convert
unsafe'' from May 2015 to the present~\footnote{This was the release date of rust 1.0}. Then we
searched on google Google Scholar using the same keywords and dates in order to to cast as wide a
net as possible.

\begin{table}
\begin{tabular}{|| p{3cm}| p{5cm} ||}
 \hline
 Database & Query \\
 \hline\hline
 ACM Digital Library & Title:(C conver* unsafe rust) AND Abstract:(C conver* unsafe rust)  \\
 IEEE Xplore & TBD \\
 Google Scholar & TBD \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

Search keywords: "C to Rust conversion", "memory-safe Rust", "legacy code migration", "unsafe blocks
mitigation", "safe programming practices".

@@

Inclusion criteria: Papers published between 2012 and 2024, written in English, focusing on
converting C code to Rust with a focus on memory safety.

Exclusion criteria: Duplicate studies, non-peer-reviewed articles, unrelated topics.

Manual search: Relevant conference proceedings, journals, and citations of selected papers.

>>>>

\documentclass[sigconf]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\usepackage[raggedrightboxes]{ragged2e}

@@

\begin{document}
\title{Rust for Operating System Development: A Systematic Literature Review}

%% \author{Shane Panter}
%%     \affiliation{%
%%     \institution{Boise State University}
%%     \city{Boise}
%%     \state{Idaho}
%%     \country{USA}}
%%     \email{shanepanter@boisestate.edu}

%% \author{Nasir U. Eisty}
%%     \affiliation{%
%%     \institution{Boise State University}
%%     \city{Boise}
%%     \state{Idaho}
%%     \country{USA}}
%%     \email{nasireisty@boisestate.edu}

@@

The migration of legacy C codebases to memory-safe programming languages like Rust is of paramount
importance in modern software development, aiming to mitigate common security vulnerabilities and
memory-related errors. This paper presents a systematic literature review (SLR) focusing on
strategies and methodologies for integrating Rust into one of the most fundamental C codebases in
the world, the Linux Kernel.  Our study aims to provide a comprehensive overview of existing
research, identify gaps, and suggest future directions in this domain. Through a rigorous search
process, we synthesized relevant studies and extracted key findings to offer insights into effective
approaches for ensuring memory safety during code migration.

@@

\subsubsection{Research Questions}
\label{sec:researchQuestions}

We have defined the following research questions to drive this study.

\begin{itemize}
    \item \textbf{RQ1:} What are the existing approaches, methodologies, and best practices for
      implementing operating system kernels in Rust?
    \item \textbf{RQ2:} What are the performance implications of using Rust for operating system
      kernel development in terms of throughput, latency, and resource utilization?
    \item \textbf{RQ3:} What are the major challenges and limitations encountered when developing
      operating system kernels in Rust, particularly in comparison to other languages?
\end{itemize}

%% Other RQ's
%% \item What are the key motivations and advantages for developing operating system kernels in Rust
%% compared to traditional languages like C or C++?
%% \item What are the fundamental design principles and features of Rust that make it suitable for
%% building operating system kernels?
%% \item What are the existing approaches, methodologies, and best practices for implementing operating
%% system kernels in Rust?
%% \item What are the major challenges and limitations encountered when developing operating system
%% kernels in Rust, particularly in comparison to other languages?
%% \item How does the safety, security, and reliability of operating system kernels developed in Rust
%% compare to those developed using other programming languages?
%% \item What are the performance implications of using Rust for operating system kernel development in
%% terms of throughput, latency, and resource utilization?
%% \item What is the current state of adoption and usage of Rust in real-world operating system kernel
%% projects, and what are the lessons learned from these experiences?
%% \item How does the Rust ecosystem support operating system kernel development, including libraries,
%% tools, and community resources?
%% \item What are the potential future directions and research opportunities for improving the development
%% of operating system kernels in Rust?
%% \item What are the implications of using Rust for operating system kernel development in terms of
%% maintenance, extensibility, and compatibility with existing hardware and software architectures?

@@

We employed a robust multi-step search strategy across three digital databases, ACM Digital Library,
IEEE Xplore, and Google Scholar in order to find all the current research that has been done. These
databases cover a wide range of current software engineering and computer science literature. We
leveraged the advanced search features of both the ACM Digital Library and IEEE Xplore to search the
titles and abstracts for keywords using boolean search operators. The ACM Digital Library and IEEE
Xplore formed the core of our keyword search and we used Google Scholar for backward and forward
snowballing.

@@

We first searched abstracts in the ACM Digital Library and metadata in IEEE Xplore with the keywords
listed in table~\ref{tab:keywords} between the years 2019~\cite{Gaynor2019-dk} and 2024. We then
used Google Scholar for forward and backward snowballing.

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & Abstract:("operating system" kernel linux) AND Abstract:(rust) & 22 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 37 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

@@

Following the guidelines outlined by Kitchenham and Charters~\cite{Stuart2007-cc} we set the
inclusion and exclusion criteria based on our research questions outlined in
section~\ref{sec:researchQuestions}. We only consider papers that are written in English, published
in conferences or journals, and within the time frame of 2019 and 2024. The published papers should
describe using the Rust programming language for either developing a new kernel, extending an
existing kernel, or authoring drivers. We will include any type of kernel architecture including
monolithic, micro, or uni as long as the paper is using the Rust programming language.
